---
title: "Penguin mass with R and Python"
author: "Taishi Huang"
date: 2026-09-23
---

This post runs Python and R in one document. Python computes mean body mass by species. R reads that table through `reticulate` and plots it.

Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/), Palmer Station Antarctica LTER. Licence: CC0.

```{r}
#| message: false
#| warning: false
library(reticulate)
library(ggplot2)

root <- Sys.getenv("QUARTO_PROJECT_ROOT")
if (!nzchar(root)) {
  root <- if (dir.exists("../../.venv")) {
    normalizePath("../..")
  } else {
    getwd()
  }
}
use_virtualenv(file.path(root, ".venv"), required = TRUE)
```

Point `reticulate` at this project's `.venv`, then compute the table in Python.

```{python}
#| message: false
#| warning: false
from palmerpenguins import load_penguins

penguins = load_penguins()
mass_by_species = (
    penguins.groupby("species", as_index=False)["body_mass_g"]
    .mean()
    .round(1)
)
mass_by_species
```

R reads the Python object with `py$mass_by_species`.

```{r}
mass_from_python <- py$mass_by_species
mass_from_python
```

```{r}
#| label: fig-mass-from-python
#| fig-cap: "Mean body mass by species. The values were computed in Python and plotted in R."
ggplot(mass_from_python, aes(x = species, y = body_mass_g)) +
  geom_col() +
  labs(x = "Species", y = "Mean body mass (g)")
```

Gentoo is the heavy species. The numbers came from pandas; the bars were drawn by ggplot2.
